### 2015-10-28 contents/title.tex
Initial ShareLaTeX Import

\title{Human Body Outline Annotation}

\author{Author\\
Imperial College London\\
180 Queen’s Gate, SW7 2AZ, London, U.K.\\
{\tt\small \{email\}@imperial.ac.uk}}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both