### 2015-11-06 contents/title.tex

\title{Estimating Correspondences of Deformable Objects ``In-the-wild"}

\author{Yuxiang Zhou, Epameinondas Antonakos, Anastasios Roussos, Stefanos Zafeiriou\\
Imperial College London\\
180 Queen’s Gate, SW7 2AZ, London, U.K.\\
{\tt\small \{yuxiang.zhou10, e.antonakos, troussos, s.zafeiriou\}@imperial.ac.uk}}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both