### 2015-11-09 contents/conclusion.tex
Final

Learning and fitting statistical deformable models (SDM) is one of the most important areas in computer vision. Generally, in order to train a SDM a set of pre-defined correspondences are required. In some objects, such as human face, semantically meaningful correspondences can be found, but require laborious human manual annotations, on other objects it is very difficult, or even impossible, to find semantically meaningful correspondences. In this paper, we propose one the first comprehensive procedures for establishing correspondences (that do not necessarily correspond to semantically meaningfull object landmarks) in arbitrary objects with minimal ammount of human annotation. We apply the proposed approach for the construction of the first, to the best of our knowledge, highly descriptive SDM for the human arm.