### 2015-11-11 contents/title.tex
Updates from ShareLaTeX

\title{Estimating Correspondences of Deformable Objects ``In-the-wild"}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both