### 2016-04-09 contents/conclusion.tex
cam ready final draft

Learning and fitting statistical deformable models (SDMs) is one of the most important areas in computer vision. Generally, in order to train a SDM, a set of predefined correspondences are required. In some objects, such as human face, semantically meaningful correspondences can be found, but require laborious manual annotations; on other objects it is very difficult, or even impossible. In this paper, we propose one of the first comprehensive procedures for establishing correspondences (that do not necessarily correspond to semantically meaningful object landmarks) in arbitrary objects with minimal amount of human annotation. We apply the proposed approach for the construction of the first, to the best of our knowledge, highly-descriptive SDM for the human arm.